---
title: "Reproducible Document"
author: "Clara"
format: html
---


```{r setup}
#| message: false
#| warning: false

library(tidyverse)

post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_small_post_meal_data <- small_post_meal_data %>%
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) %>%
  filter(glykemi == 0) %>%
  select(-glykemi)

tidied_post_meal_data <- tidier_small_post_meal_data %>%
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First degree relative")
  )
```

## Writing R code

2 + 2 

```{r}
2 + 2
```
## About me 

- Clara Solà
- Steno Diabetes Center Copenhagen

Currently doing a **PhD** on *female health, exercise and T1D*

## Simple Code

```{r}
3 * 3
```
## Testing for Git

Its very cold

## Showing the data

```{r}
post_meal_data

glimpse(post_meal_data)
```
## Plot one continuos variable 

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

See @fig-bmi-histo

## Exercise discrete plots

```{r}
#| fig-cap: " Group "
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar

## Plotting two discrete variables 

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Side by side plots 

```{r fig-bmi-glycemia}
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#|     - "Distribution of BMI."
#|     - "Number of those with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```


## Plotting two continuos variables 

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Selecting data

```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```

## Renaming columns

```{r}
small_post_meal_data %>%
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Select and rename tasks

```{r}
select(post_meal_data, contains("OGTT"))
```

```{r}
post_meal_data %>%
  select(contains(".."), -contains("..."))
```

```{r}
post_meal_data %>%
  rename(
    Glucose.minus5 = P.Glucose..5.OGTT,
    Insulin.minus5 = Insulin..5.OGTT.X
  )
```



## Filtering data

```{r}
small_post_meal_data %>%
  filter(glykemi == 1)

small_post_meal_data %>%
  filter(BMI >= 25)

small_post_meal_data %>%
  filter(Group == "CTR")

small_post_meal_data %>%
  filter(Group == "CTR" | BMI >= 25)
```

## Modifying data 

```{r}
tidier_small_post_meal_data %>%
  mutate(insulin_glucose_ratio = auc_insulin / auc_glucose)

tidier_small_post_meal_data %>%
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First degree relative")
  )
```

## Practice with filter and mutate

```{r}
post_meal_data %>%
  filter(BMI >= 25 & BMI <= 40 & Group == "FDR")
```
```{r}
post_meal_data %>%
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100
  ) %>%
  filter(percent_body_fat > 30)
```

## Visualizing with a jitter plot

```{r}
tidied_post_meal_data %>%
  mutate(overweight = ifelse(BMI >= 25, "Overweight", "Not overweight")) %>%
  ggplot(aes(x = overweight, y = insulin_glucose_ratio)) +
  geom_jitter()
```

## Plotting three or more variables 


```{r}
tidied_post_meal_data %>%
  filter(BMI < 30) %>%
  mutate(young = if_else(Age < 40, "Young", "Old")) %>%
  ggplot(aes(x = BMI, y = insulin_glucose_ratio, colour = Group)) +
  geom_point() +
  facet_grid(rows = vars(young)) +
  labs(
    x = "Body Mass Index (BMI)",
    y = "Insulin to glucose ratio"
  )
```

## Make multi-dimensional plots 

```{r}
#| label: fig-multidimensional
#| fig-cap: "Two, mult-dimensional plots of our data."
#| fig-subcap:
#|  - "Distribution of percent body fat by group and age category."
#|  - "Relationship between BMI and bone mineral density, split by insulin sensitivity."
#|layout-ncol: 2

# Task 1.
post_meal_data %>%
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100,
    young = if_else(Age < 40, "Young", "Old")
  ) %>%
  ggplot(aes(x = Group, y = percent_body_fat, colour = young)) +
  geom_jitter(width = 0.2)
```

```{r}
#| label: fig-multidimensional2
#| fig-cap: "Two, mult-dimensional plots of our data."
#| fig-subcap:
#|  - "Distribution of percent body fat by group and age category."
#|  - "Relationship between BMI and bone mineral density, split by insulin sensitivity."
#|layout-ncol: 2

post_meal_data %>%
  mutate(
    insulin_glucose_ratio = (auc_ins / auc_pg),
    insulin_insensitive = if_else(insulin_glucose_ratio > mean(insulin_glucose_ratio), "Insensitive", "Sensitive")
  ) %>%
  ggplot(aes(x = BMI, y = Bone.mineral.DXA, colour = insulin_insensitive)) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(insulin_insensitive))
```


## Calculating summary statistics

```{r}
tidied_post_meal_data %>%
  summarise(
    max_bmi = max(BMI),
    min_bmi = min(BMI)
  )
```
## Summarise with group by

```{r}
#| label: tbl-mean-age-bmi
#| tbl-cap: "Mean values of Age and BMI for each group"

tidied_post_meal_data %>%
  group_by(Group) %>%
  summarise(
    "Mean Age (yrs)" = round (mean(Age) , 1),
    "Mean BMI (kg/m^2^)" = round (mean(BMI) , 1) 
  ) %>%
  ungroup() %>% 
  knitr::kable()

```
