---
title: "Reproducible Document"
author: "Clara"
format: html
---


```{r setup}
#| message: false
#| warning: false

library(tidyverse)

post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)
```

## Writing R code

2 + 2 

```{r}
2 + 2
```
## About me 

- Clara Solà
- Steno Diabetes Center Copenhagen

Currently doing a **PhD** on *female health, exercise and T1D*

## Simple Code

```{r}
3 * 3
```
## Testing for Git

Its very cold

## Showing the data

```{r}
post_meal_data

glimpse(post_meal_data)
```
## Plot one continuos variable 

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

See @fig-bmi-histo

## Exercise discrete plots

```{r}
#| fig-cap: " Group "
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar

## Plotting two discrete variables 

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Side by side plots 

```{r fig-bmi-glycemia}
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#|     - "Distribution of BMI."
#|     - "Number of those with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```


## Plotting two continuos variables 

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Selecting data

```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```

## Renaming columns

```{r}
small_post_meal_data %>%
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```
